`flippant` -- An `R` package for the automated  analysis of dithionite

### 2015-09-18 inst/manuscript/manuscript.rmd
Integrate changes by Birgit;

# Introduction
Scramblases are proteins mediating the mass-action driven equilibration of 
lipids between the leaflets of lipid bilayers that constitute biological membranes
[@waechter_flippases_2013]. Independent of energy equivalents such as ATP they 
facilitate this ``flipping'' by providing to polar lipid head groups a mechanism
of transition through the hydrophobic environment of membranes, thus reducing 
the massive energetic cost that process would carry otherwise 
[@kornberg_inside-outside_1971]. While the underlying molecular mechanisms in 
many cases remain enigmatic [for a discussion including exceptions see @ernst_phospholipid_2015], and molecules supporting flipping processes as 
central as the equilibration of the ubiquitous phosphoglycerolipids across the 
membrane of the endoplasmic reticulum, where they are synthesized in the 
cytoplasmic monolayer, have yet to be identified **(I was looking for a recent 
citation that clearly spelled that out - in vain. Do you have a proposal?)**, 
there has been much recent progress in scramblase identification and 
characterization. Molecules shown to have scramblase activity include 
opsin/rhodopsin [@menon_opsin_2011;@goren_constitutive_2014], TMEM16 
[@malvezzi_ca2-dependent_2013], $\beta$ 2-adrenergic and adenosine A2A 
receptors [both @goren_constitutive_2014]. **Would be good to include/replace some 
with non-in-house identifications/characterizations. What else is out there?**

To interrogate factors of this functional class, we have developed a biochemical 
assay suited to characterize scramblase activity in purified proteins and protein 
extracts [@chang_chemical_2004] based on the reconstitution of scramblases or 
candidate proteins into synthetic liposomes with a lipid
composition that includes nitrobenzoxadiazole (NBD) labeled fluorescent lipids.
Upon exposure of these liposomes to membrane-impermeable reducing agents such as 
dithionite, fluorophores in the external leaflet are quenched and fluorescence in scramblase-free liposomes accordingly quickly drops to approximately 50% of the 
initial value. However, for proteoliposomes reconstituted with proteins 
conferring a scramblase activity, fluorescence disappears entirely, as
fast scramblase-mediated equilibration between outer (exposed) and inner
(protected) leaflet of the bilayer renders all fluorescent lipids
accessible to reduction. By titration of the protein amount reconstituted into
the liposome membranes it is thus possible to characterize scramblases and probe
their activity. Beyond the identification of scramblase activities in crude 
membrane preparations [@chang_chemical_2004;@rajasekharan_flip-flop_2011] and 
functional validation of candidate scramblases [@menon_opsin_2011;@malvezzi_ca2-dependent_2013;@goren_constitutive_2014], the 
assay has also been used to characterize the stoichiometry of scramblase 
homo-multimers reconstituting into the proteoliposomes 
[@ploier_perturbed_submitted] and probe the ATP-dependence of flippases, 
molecules that under energy-consumption flip lipids against a concentration 
gradient between bilayer leaflets [@soupene_atp8a1_2008].
**Are you aware of further use? More adoption would fortify the need for the 
package ...**

>>>>

# Introduction
Scramblases are proteins mediating the mass-action driven equilibration of 
lipids between the leaflets of lipid bilayers that constitute biological membranes
[@waechter_flippases_2013]. Independent of energy equivalents such as ATP they 
facilitate this ``flipping'' by providing to polar lipid head groups a mechanism
of transition through the hydrophobic environment of membranes, thus reducing 
the massive energetic cost that process would carry otherwise 
[@kornberg_inside-outside_1971]. The underlying molecular mechanisms remain 
largely enigmatic [for a discussion including exceptions see @ernst_phospholipid_2015].
Proteins supporting flipping processes as central as the equilibration of the 
ubiquitous phosphoglycerolipids across the membrane of the endoplasmic 
reticulum, where they are synthesized in the cytoplasmic monolayer, have yet to 
be identified **(I was looking for a recent 
citation that clearly spelled that out - in vain. Do you have a proposal?)**. 
None the less, there has been much recent progress in scramblase identification 
and characterization. Proteins shown to have scramblase activity include 
FtsW [@mohammadi_identification_2011], opsin/rhodopsin
[@menon_opsin_2011;@goren_constitutive_2014], TMEM16 
[@malvezzi_ca2-dependent_2013], $\beta$ 2-adrenergic and adenosine A2A 
receptors [both @goren_constitutive_2014]. Extrapolating from the latter findings 
and the opsin/rhodopsin case, we speculate that lipid scrambling may, in fact, 
be an intrinsic property of G-protein coupled receptors in general 
[@goren_constitutive_2014].

To interrogate factors of this functional class, we have developed a biochemical 
assay suited to characterize scramblase activity of purified proteins and protein 
extracts [@chang_chemical_2004] based on the reconstitution of scramblases or 
candidate proteins into preformed synthetic liposomes that contain trace amounts
of nitrobenzoxadiazole (NBD) labeled fluorescent lipids.
Upon addition of membrane-impermeable reducing agents such as dithionite to 
these liposomes, fluorophores in the external leaflet are quenched and fluorescence in scramblase-free liposomes accordingly quickly drops to approximately 50% of the 
initial value. However, for proteoliposomes reconstituted with proteins 
conferring a scramblase activity, fluorescence disappears entirely, as
fast scramblase-mediated equilibration between outer (exposed) and inner
(protected) leaflet of the bilayer renders all fluorescent lipids
accessible to reduction. By titration of the protein amount reconstituted into
the liposome membranes it is thus possible to characterize scramblases and probe
their activity. Beyond the identification of scramblase activities in crude 
membrane preparations [@chang_chemical_2004;@rajasekharan_flip-flop_2011] and 
functional validation of candidate scramblases [@menon_opsin_2011;@malvezzi_ca2-dependent_2013;@goren_constitutive_2014], the 
assay has also been used to characterize the stoichiometry of scramblase 
homo-multimers reconstituting into the proteoliposomes 
[@ploier_perturbed_submitted] and probe the ATP-dependence of flippases, 
molecules that under energy-consumption flip lipids against a concentration 
gradient between bilayer leaflets [@soupene_atp8a1_2008].
**Are you aware of further use? More adoption would fortify the need for the 
package ...**
